---
title: "Lobster Report"
Owner: Victor Knowles
Collaborator: Ramit Guha
format: html
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
```

## Citation

Santa Barbara Coastal LTER, D. Reed, and R. Miller. 2022. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/25aa371650a671bafad64dd25a39ee18 (Accessed 2026-04-14).

## Lobster Abundance & Size File

```{r}
lobster_abundance <- read_csv(here::here(
  "Data Folder",
  "Time-series of lobster abundance and size",
  "Lobster_Abundance_All_Years_20220829.csv"
))
```

```{r}
lobster_abundance <- lobster_abundance %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))
```

```{r}
# `group_by() %>% summarize()` practice
max_lobster_traps <- lobster_traps %>% 
    group_by(SITE, MONTH) %>%
    summarize(MAX_TRAPS = max(TRAPS, na.rm = TRUE))
```

## Lobster Trap Counts

```{r}
here::here(
  "Data Folder",
  "Time-series of lobster trap buoy counts",
  "Lobster_Trap_Counts_All_Years_20210519.csv"
)
```

```{r}
lobster_traps <- read_csv(
  here::here(
    "Data Folder",
    "Time-series of lobster trap buoy counts",
    "Lobster_Trap_Counts_All_Years_20210519.csv"
  )
)
```

```{r}
lobster_traps <- lobster_traps %>% 
    mutate(TRAPS = na_if(TRAPS, -99999))
```

```{r}
not_napl <- lobster_traps %>% 
    filter(SITE != "NAPL")
```

```{r}
carp_20_traps <- lobster_traps %>% 
    filter(SITE == "CARP" & TRAPS > 20)
```

```{r}
# `group_by() %>% summarize()` practice
max_lobster_traps <- lobster_traps %>% 
    group_by(SITE, MONTH) %>%
    summarize(MAX_TRAPS = max(TRAPS, na.rm = TRUE))
```

## Summary of Abstract

This study tracks California spiny lobster populations along the Santa Barbara Channel to understand how fishing impacts kelp forest ecosystems. Researchers collect data on lobster abundance, size, and fishing pressure, especially comparing sites inside marine protected areas (MPAs) to those outside. The dataset includes annual diver surveys of lobster populations at five reef sites and regular measurements of fishing activity (via trap counts) during the fishing season across multiple locations.

```{r}
ggplot(data = lobster_abundance,
       aes(x = SIZE_MM)) +
  geom_histogram() +
  facet_wrap(~SITE)
```

```{r}
## Run this chunk to test exercises ##

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)

lobster_traps <- read_csv(
  here::here(
    "Data Folder",
    "Time-series of lobster trap buoy counts",
    "Lobster_Trap_Counts_All_Years_20210519.csv"
  )
) %>%
  mutate(TRAPS = na_if(TRAPS, -99999))
```

```{r}
# line and point plot
ggplot(data = lobsters_traps_summarize, aes(x = YEAR, y = TOTAL_TRAPS)) +
    geom_point(aes(color = SITE)) +
    geom_line(aes(color = SITE))

```
